### 2023-04-23 course/csic5011/2023/project2/project2.tex
project2

@@

\lecture{Final Project}{Yuan Yao}{Due: 23:59 Sunday 19 May, 2019}{3 May, 2019}

@@

\begin{enumerate}
\item Pick up ONE (or more if you like) favourite dataset below to work. If you would like to work on a different problem outside the candidates we proposed, please email course instructor about your proposal.  
\item Team work: we encourage you to form small team, up to THREE persons per group, to work on the same problem. Each team must submit:
\subitem(a) \emph{ONE report}, \emph{with a clear remark on each person's contribution}. The report can be in the format of a \emph{technical report within 8 pages}, e.g. NIPS conference style 
\begin{center}
\url{https://nips.cc/Conferences/2016/PaperInformation/StyleFiles} 
\end{center}
or of a \emph{poster}, e.g. 
\begin{center}%\url{http://math.stanford.edu/~yuany/publications/poster_CleaveBioCPH2017_ForReview.pptx}
\url{https://github.com/yuany-pku/2017_math6380/blob/master/project1/DongLoXia_poster.pptx}
\end{center}
\subitem(b) \emph{ONE short presentation video within 10 mins}, e.g. in Youtube link. You may submit your presentation slides together with the video link to help understanding. 
\item In the report, (1) design or raise your scientific problems (a good problem is often more important than solving it); (2) show your main results with a careful analysis supporting the results toward answering your problems. Remember: scientific analysis and reasoning are more important than merely the performance results. Source codes may be submitted through email as a zip file, or as an appendix if it is not large.    
\item Submit your report by email or paper version no later than the deadline, to the following address (\href{mailto:datascience.hw@gmail.com}{datascience.hw@gmail.com}) with Title: \underline{CSIC 5011: Project 2}. % (\href{mailto:datascience\_hw@126.com}{datascience\_hw@126.com}). 
\end{enumerate}

@@

\section*{Open Peer Review}

In this exercise of open peer review, please write down your comments of the \emph{reports rather than of your own team} in the following format. Be considerate and careful with a precise description, avoiding offensive language.

Deadline is 23:59 May 25, 2019. Submit your review in plain text to the email address (\href{mailto:datascience.hw@gmail.com}{datascience.hw@gmail.com}) with Title: \underline{CSIC 5011: Project 2 Review}. Rebuttal is open afterwards.

\begin{itemize}
\item Summary of the report.
\item Describe the strengths of the report. 
\item Describe the weaknesses of the report.
\item Evaluation on quality of writing (1-5): Is the report clearly written? Is there a good use of examples and figures? Is it well organized? Are there problems with style and grammar? Are there issues with typos, formatting, references, etc.? Please make suggestions to improve the clarity of the paper, and provide details of typos.
\item Evaluation on presentation (1-5): Is the presentation clear and well organized? Are the language flow fluent and persuasive? Are the slides clear and well elaborated? Please make suggestions to improve the presentation.
\item Evaluation on creativity (1-5): Does the work propose any genuinely new ideas? Is this a work that you are eager to read and cite? Does it contain some state-of-the-art results? As a reviewer you should try to assess whether the ideas are truly new and creative. Novel combinations, adaptations or extensions of existing ideas are also valuable.
\item Confidence on your assessment (1-3)
(3- I have carefully read the paper and checked the results, 2- I just browse the paper without checking the details, 1- My assessment can be wrong)
\end{itemize}

\section*{Rebuttal}
The rebuttal period starts from now, till 23:59 May 31, 2019. Restrict the number of characters of your rebuttal within {\bf{5,000}}. Submit your rebuttal in \emph{PLAIN TEXT} or \emph{Word Document} format to the email address (\href{mailto:datascience.hw@gmail.com}{datascience.hw@gmail.com}) with Title: \underline{CSIC 5011: Project 2 Rebuttal}.

The following tips of rebuttal might be helpful for you to follow:

1. The main aim of the rebuttal is to answer any specific questions that the reviewers might have raised, or to clarify any misunderstanding of the technical content of the paper.

2. Keep your rebuttal short, to-the-point, and specific. In our experience, such rebuttals have the maximum impact.

3. Always be polite and professional. Refrain from name calling or rude comments, especially in response to negative reviews.

4. Highlight the changes in your manuscripts had you made a simple revision.

@@

\section{Order the faces by Diffusion Map}
The following dataset contains 33 faces of the same person ($Y\in \R^{112\times 92\times33}$) in different angles,

You may create a data matrix $X\in \R^{n\times p}$ where $n=33,p=112\times92=10304$ \\ (e.g. {\texttt{X=reshape(Y,[10304,33])';}} in matlab).

\begin{enumerate}
\item Explore the Diffusion map, or the second smallest eigenvector of Markov Chains defined on the point cloud data, to order the faces, i.e., let $W_{ij}=\exp(-\|x_i - x_j\|^2/t)$ with $D=\diag(\sum_j W_{ij})$ and define $L=D^{-1} W - I$, clearly $\lambda_0=0$ and take the (second) smallest nonzero eigenvalue $\lambda_1$ with corresponding eigenvector $v_1$, sort the faces by values $v_1(i)$, $i=1,\ldots,n$.  
\item Explore the MDS-embedding of the 33 faces on top two eigenvectors: order the faces according to the top 1st eigenvector and visualize your results with figures. 
\item Explore the ISOMAP-embedding of the 33 faces on the $k=5$ nearest neighbor graph and compare it against the MDS results. Note: you may try Tenenbaum's Matlab code \\
{\url{https://github.com/yao-lab/yao-lab.github.io/blob/master/data/isomapII.m}}
\item Explore the LLE-embedding of the 33 faces on the $k=5$ nearest neighbor graph and compare it against ISOMAP. Note: you may try the following Matlab code \\
{\url{https://github.com/yao-lab/yao-lab.github.io/blob/master/data/lle.m}}
\end{enumerate}

You might explore larger datasets with other manifold learning methods, for example the PubFig dataset et al.

@@

As shown in Figure \ref{fig:karate}, node 1 represents the coach of the club and node 34 is the owner (president) of the club. The undirected, unweighted edges between nodes represent
the affinity relation between club members. The story behind the network is this: the coach would like to raise the instruction fee while the president does not allow this; the conflicts finally result in a fission of the club -- the coach leaves the club with his funs and sets up his own club marked in red, and the blue nodes remain in the old club with the president.  
\begin{figure}[h]
\centering
\includegraphics[width=0.3\textwidth]{./figures/karate1-a_old.eps}
\caption{Zachery's Karate Club Network}\label{fig:karate}
\end{figure}

@@

\section{Identification of Raphael's paintings from the forgeries}

The following data, provided by Prof. Yang WANG from HKUST,

\url{https://drive.google.com/folderview?id=0B-yDtwSjhaSCZ2FqN3AxQ3NJNTA&usp=sharing}

\noindent contains a 28 digital paintings of Raphael or forgeries. Note that there are both jpeg and tiff files, so be careful with the bit depth in digitization. The following file

\url{https://docs.google.com/document/d/1tMaaSIrYwNFZZ2cEJdx1DfFscIfERd5Dp2U7K1ekjTI/edit}

\noindent contains the labels of such paintings, which are 
\begin{enumerate}
\item[1] Maybe Raphael - Disputed
\item[2] Raphael
\item[3] Raphael
\item[4] Raphael
\item[5] Raphael
\item[6] Raphael
\item[7] Maybe Raphael - Disputed
\item[8] Raphael
\item[9] Raphael
\item[10] Maybe Raphael - Disputed
\item[11] Not Raphael
\item[12] Not Raphael
\item[13] Not Raphael
\item[14] Not Raphael
\item[15] Not Raphael
\item[16] Not Raphael
\item[17] Not Raphael
\item[18] Not Raphael
\item[19] Not Raphael
\item[20] My Drawing (Raphael?)
\item[21] Raphael
\item[22] Raphael
\item[23] Maybe Raphael - Disputed
\item[24] Raphael
\item[25] Maybe Raphael - Disputed
\item[26] Maybe Raphael - Disputed
\item[27] Raphael
\item[28] Raphael
\end{enumerate}
Can you exploit the known Raphael vs. Not Raphael data to predict the identity of those 6 disputed paintings (maybe Raphael)? The following student poster report seems a good exploration

\url{https://yao-lab.github.io/2015.fall.pku/poster/Raphael_LI%2CYue_1300010601.pdf}

The following paper by Haixia Liu, Raymond Chan, and me studies Van Gogh's paintings which might be a reference for you:

\url{http://dx.doi.org/10.1016/j.acha.2015.11.005}

@@

{\it All the following datasets have been given in project 1. But you may explore these data with various new techniques learned in class, e.g. Robust PCA, Sparse PCA, Robust PCA by GAN (\url{https://github.com/zhuwzh/Robust-GAN-Scatter}), Manifold Learning, Topological Data Analysis (Mapper, Persistent Homology, etc.). }

@@

\section{SNPs Data}
 This dataset contains a data matrix $X\in \R^{p\times n}$ of about $n=650,000$ columns of SNPs (Single Nucleid Polymorphisms) and $p=1064$ rows of peoples around the world. Each element is of three choices, $0$ (for `AA'), $1$ (for `AC'), $2$ (for `CC'), and some missing values marked by $9$.

\url{https://www.dropbox.com/l/scl/AADN80paNFy1yB5gyYzNVOfkZGj9SiVDlZo}
%\url{http://math.stanford.edu/~yuany/course/ceph_hgdp_minor_code_XNA.txt.zip}

\noindent which is big (151MB in zip and 2GB original txt). Moreover, the following file contains the region where each people comes from, as well as two variables {\texttt{ind1}} and{\texttt{ind2}} such that $X({\texttt{ind1}},{\texttt{ind2}})$ removes all missing values.

%\url{http://math.stanford.edu/~yuany/course/data/HGDP_region.mat}
\url{https://yao-lab.github.io/data/HGDP_region.mat}

\noindent More detailed information about these persons in the dataset can be also found at

\url{https://yao-lab.github.io/data/HGDPid_populations_ALL.xls}

Some results by PCA can be found in the following paper, Supplementary Information.

@@

@@

When $\theta=\infty$ without noise, classical MDS will work; but for a finite $\theta$ with noisy measurements, SDP approach can be useful. You may try the matlab package SNLSDP by Kim-Chuan Toh, Pratik Biswas, and Yinyu Ye, downladable at \url{http://www.math.nus.edu.sg/~mattohkc/SNLSDP.html}.

%Attention: this last dataset is relatively big with about 2GB size. 
%
%You can login my server:
%
%\texttt{ssh einstein@162.105.205.92}
%
%\noindent using the password I provided on class. On the read only folder \texttt{/data/snp/}, you will find all the data in both .txt and .mat (\texttt{data.mat, HGDP\_region.mat, readme.m}).

%\subsection{Bird Flu Dataset} (courtesy of Steve Smale and Cissy) This dataset 162 H5N1 (bird flu) virus sequences discovered around the world:
%
%\url{http://www.math.pku.edu.cn/teachers/yaoy/data/birdflu_seq162.txt} 
%
%Locations of such virus discovered are reported with latitude and longitude coordinates on the globe:
%
%\url{http://www.math.pku.edu.cn/teachers/yaoy/data/birdflu_latgrat.txt} 
%
%Pairwise geodesic distances between these 162 sites are constructed as  
%
%\url{http://www.math.pku.edu.cn/teachers/yaoy/data/birdflu_geodist.txt}
%
%A kernel-induced $l_2$-distances between 162 virus sequences are given in 
%
%\url{http://www.math.pku.edu.cn/teachers/yaoy/data/birdflu_l2dist.txt}
\end{document}

>>>>

\graphicspath{{figures/}{../book.datasci/figures/}{../mySlides/figures/}}

@@

\lecture{Final Project}{Yuan Yao}{Due: 23:59 18 May, 2023}{24 Apr, 2023}

@@

\begin{enumerate}
\item Pick up ONE (or more if you like) favourite dataset below to work. If you would like to work on a different problem outside the candidates we proposed, please email course instructor about your proposal.  
\item Team work: we encourage you to form small team, up to FOUR persons per group, to work on the same problem. Each team must submit:
\subitem(a) \emph{ONE report}, \emph{with a clear remark on each person's contribution}. The report can be in the format of a \emph{technical report within 8 pages}, e.g. NIPS conference style 
\begin{center}
\url{https://nips.cc/Conferences/2016/PaperInformation/StyleFiles} \\
\noindent and a sample file at \\
\url{https://arxiv.org/pdf/1606.04930.pdf} 
\end{center}
or of a \emph{poster}, e.g. 
\begin{center}%\url{http://math.stanford.edu/~yuany/publications/poster_CleaveBioCPH2017_ForReview.pptx}
\url{https://github.com/yuany-pku/2017_math6380/blob/master/project1/DongLoXia_poster.pptx}
\end{center}
\subitem(b) \emph{ONE short presentation video within 10 mins}, e.g. in Youtube or Bilibili link. You may submit your presentation slides together with the video link to help understanding. 
\item In the report, (1) design or raise your scientific problems (a good problem is sometimes more important than solving it); (2) show your main results with a careful analysis supporting the results toward answering your problems. Remember: scientific analysis and reasoning are more important than merely the performance results. Reproducible source codes may be submitted via a \url{https://github.com/} link, through email as a zip file, or as an appendix of the report if it is not large. 
\item Upon the start, submit your grouping information (representer/member) to the following address (\href{mailto:datascience.hw@gmail.com}{datascience.hw@gmail.com}) with Title: \underline{CSIC 5011: Project 2 [grouping]}. Then we shall create public folders and you should submit your report to your group folder under subfolder \texttt{[GROUP]/report}. % (\href{mailto:datascience\_hw@126.com}{datascience\_hw@126.com}). 
\end{enumerate}

@@

%\section*{Open Peer Review}
%
%In this exercise of open peer review, please write down your comments of the \emph{reports rather than of your own team} in the following format. Be considerate and careful with a precise description, avoiding offensive language. 
%
%Deadline is 23:59 Sunday May 31, 2020. Submit your review in \textbf{PLAIN TEXT} to the corresponding group's subfolder \texttt{0.peer-review/[GROUP]}. 
%% email address (\href{mailto:datascience.hw@gmail.com}{datascience.hw@gmail.com}) with Title: \underline{CSIC 5011: Project 2 Review}. Rebuttal is open afterwards.
%
%\begin{itemize}
%\item Summary of the report.
%\item Describe the strengths of the report. 
%\item Describe the weaknesses of the report.
%\item Evaluation on quality of writing (1-5): Is the report clearly written? Is there a good use of examples and figures? Is it well organized? Are there problems with style and grammar? Are there issues with typos, formatting, references, etc.? Please make suggestions to improve the clarity of the paper, and provide details of typos.
%\item Evaluation on presentation (1-5): Is the presentation clear and well organized? Are the language flow fluent and persuasive? Are the slides clear and well elaborated? Please make suggestions to improve the presentation.
%\item Evaluation on creativity (1-5): Does the work propose any genuinely new ideas? Is this a work that you are eager to read and cite? Does it contain some state-of-the-art results? As a reviewer you should try to assess whether the ideas are truly new and creative. Novel combinations, adaptations or extensions of existing ideas are also valuable.
%\item Confidence on your assessment (1-3)
%(3- I have carefully read the paper and checked the results, 2- I just browse the paper without checking the details, 1- My assessment can be wrong)
%\end{itemize}
%
%\newpage
%
%\section*{Rebuttal}
%The rebuttal period starts from now, till 23:59 June 5, 2020. Restrict the number of characters of your rebuttal within {\bf{5,000}}. Submit your rebuttal in \textbf{PLAIN TEXT} format to the 
%your group folder under subfolder \texttt{[GROUP]/rebuttal}. 
%%email address (\href{mailto:datascience.hw@gmail.com}{datascience.hw@gmail.com}) with Title: \underline{CSIC 5011: Project 2 Rebuttal}.
%
%The following tips of rebuttal might be helpful for you to follow:
%
%1. The main aim of the rebuttal is to answer any specific questions that the reviewers might have raised, or to clarify any misunderstanding of the technical content of the paper.
%
%2. Keep your rebuttal short, to-the-point, and specific. In our experience, such rebuttals have the maximum impact.
%
%3. Always be polite and professional. Refrain from name calling or rude comments, especially in response to negative reviews
%
%4. Highlight the changes in your manuscripts had you made a simple revision.
%
%\newpage

\section{Human Prefrontal Cortex Development Data}
This dataset contains a single cell gene expression matrix $X\in \R^{n\times p}$ of $n=24153$ genes and $p=2394$ cells. Each value is in the unit of transcript-per-million ({TPM}\href{https://www.rna-seqblog.com/rpkm-fpkm-and-tpm-clearly-explained/}). The file of the dataset is in tab-delimited text format, where each row represents one gene (the first row is the cell ID) and each column represents one cell (the first column is the gene name). The size is about 33.2 Mb in gzipped format and about 160 Mb after decompression. Link to download the data:

\url{https://www.ncbi.nlm.nih.gov/geo/download/?acc=GSE104276&format=file&file=GSE104276%5Fall%5Fpfc%5F2394%5FUMI%5FTPM%5FNOERCC%2Exls%2Egz}

The single cells were collected in the human embryonic prefrontal cortex (PFC) from gestational weeks (GW) 8 to 26. More specifically, the cells were collected at GW8, GW9, GW10, GW12, GW13, GW16, GW19, GW23 and GW26, and included three and two biological replicates at GW10 and GW23, respectively. In the gene expression matrix, the cell ID contains both the time point and the donor information. For example, cell "GW10\_PFC1\_sc35" indicates that the cell was collected at gestational week 10 from donor 1. To learn more about the dataset you can visit the Gene Expression Omnibus (\href{https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE104276}{GSE104276}) or read the original publication in \textit{Nature}:

\begin{itemize}
\item 
Zhong, S., Zhang, S., Fan, X. et al. A single-cell RNA-seq survey of the developmental landscape of the human prefrontal cortex, \emph{Nature} {\bf 555}, 524-528 (2018). \\
\href{https://www.nature.com/articles/nature25980}{https://doi.org/10.1038/nature25980}.
\end{itemize}

Can you identify subgroups of cells from the data, and trace the developmental trajectories of these cells? Dimensionality reduction methods and single cell topological data analysis methods may help.

\section{Dynamic Simplicial Complexes of Teacher-Student Contacts}
Infection diseases like COVID-19 now influenced everyone's life and are a challenge for human to fight. The following data set contains the temporal network of contacts between the children and teachers in a primary school used in the study published in BMC Infectious Diseases 2014, 14:695.

\url{http://www.sociopatterns.org/datasets/primary-school-temporal-network-data/}

The file contains a tab-separated list representing the active contacts during 20-second intervals of the data collection. Each line has the form ``\texttt{t i j Ci Cj}", where \texttt{i} and \texttt{j} are the anonymous IDs of the persons in contact, \texttt{Ci} and \texttt{Cj} are their classes, and the interval during which this contact was active is $[ \texttt{t} - 20s, \texttt{t} ]$. If multiple contacts are active in a given interval, you will see multiple lines starting with the same value of \texttt{t}. Time is measured in seconds.

Based on such data, simplices are extracted through cliques of simultaneous contacts in the following directory by Austin Benson et al.:

\url{https://github.com/arbenson/ScHoLP-Data/tree/master/contact-primary-school}

\noindent Specifically, for every unique timestamp in the dataset, a simplex is constructed for every maximal
clique amongst the contact edges that exist for that timestamp. Timestamps were
recorded in 20 second intervals.

You may construct simplicial complexes (a collection of simplices that is closed under inclusion) from such data and study its topological properties (e.g. Betti numbers) with dynamic changes.

For those who are studying genomics can explore the phylogenetic trees and networks for genomics data of hCoV-19 at:

\url{https://www.gisaid.org/}

@@

%\section{Robust Mean and Covariance Estimate by GAN}
%
%Choose a dataset in your favor, e.g. the stock price data from Yahoo!Finance shown in the notebook
%
%\url{https://github.com/yao-lab/yao-lab.github.io/blob/master/data/RobustGAN/Robust_GAN.ipynb}
%
%\noindent or other datasets in this final projects. Try:
%\begin{enumerate}
%\item robust mean and covariance (scatter) matrix estimation using GAN with different models (Gaussian, student-$t$, and general Elliptical distributions);
%\item outlier detection using distribution of discriminator values;
%\item robust PCA using top eigenvectors of robust covariance (scatter) matrices. 
%\end{enumerate}

@@

As shown in Figure \ref{fig:karate}, node 1 represents the coach of the club and node 34 is the owner (president) of the club. The undirected, unweighted edges between nodes represent
the affinity relation between club members. The story behind the network is this: the coach would like to raise the instruction fee while the president does not allow this; the conflicts finally result in a fission of the club -- the coach leaves the club with his funs and sets up his own club marked in red, and the blue nodes remain in the old club with the president.  
\begin{figure}[h]
\centering
\includegraphics[width=0.3\textwidth]{karate1-a_old.eps}
\caption{Zachery's Karate Club Network}\label{fig:karate}
\end{figure}

@@

%\section{Order the faces by Diffusion Map}
%The following dataset contains 33 faces of the same person ($Y\in \R^{112\times 92\times33}$) in different angles,
%
%\url{https://github.com/yao-lab/yao-lab.github.io/blob/master/data/face.mat}
%
%You may create a data matrix $X\in \R^{n\times p}$ where $n=33,p=112\times92=10304$ \\ (e.g. {\texttt{X=reshape(Y,[10304,33])';}} in matlab).
%
%\begin{enumerate}
%\item Explore the Diffusion map, or the second smallest eigenvector of Markov Chains defined on the point cloud data, to order the faces, i.e., let $W_{ij}=\exp(-\|x_i - x_j\|^2/t)$ with $D=\diag(\sum_j W_{ij})$ and define $L=D^{-1} W - I$, clearly $\lambda_0=0$ and take the (second) smallest nonzero eigenvalue $\lambda_1$ with corresponding eigenvector $v_1$, sort the faces by values $v_1(i)$, $i=1,\ldots,n$.  
%\item Explore the MDS-embedding of the 33 faces on top two eigenvectors: order the faces according to the top 1st eigenvector and visualize your results with figures. 
%\item Explore the ISOMAP-embedding of the 33 faces on the $k=5$ nearest neighbor graph and compare it against the MDS results. Note: you may try Tenenbaum's Matlab code \\
%{\url{https://github.com/yao-lab/yao-lab.github.io/blob/master/data/isomapII.m}}
%\item Explore the LLE-embedding of the 33 faces on the $k=5$ nearest neighbor graph and compare it against ISOMAP. Note: you may try the following Matlab code \\
%{\url{https://github.com/yao-lab/yao-lab.github.io/blob/master/data/lle.m}}
%\end{enumerate}
%
%You might explore larger datasets with other manifold learning methods, for example the PubFig dataset et al.
%
%\url{http://www.cs.columbia.edu/CAVE/databases/pubfig/}
%

\section{Some classical image datasets}

\subsection{Face dataset}

The following dataset contains 33 faces of the same person ($Y\in \R^{112\times 92\times33}$) in different angles,

You may create a data matrix $X\in \R^{n\times p}$ where $n=33,p=112\times92=10304$ \\ (e.g. {\texttt{X=reshape(Y,[10304,33])';}} in matlab). You may try Python package

\url{https://scikit-learn.org/stable/modules/manifold.html#}

\noindent For example, 
\begin{enumerate}
\item Explore the Diffusion map, or the second smallest eigenvector of Markov Chains defined on the point cloud data, to order the faces, i.e., let $W_{ij}=\exp(-\|x_i - x_j\|^2/t)$ with $D=\diag(\sum_j W_{ij})$ and define $L=D^{-1} W - I$, clearly $\lambda_0=0$ and take the (second) smallest nonzero eigenvalue $\lambda_1$ with corresponding eigenvector $v_1$, sort the faces by values $v_1(i)$, $i=1,\ldots,n$.  
\item Explore the MDS-embedding of the 33 faces on top two eigenvectors: order the faces according to the top 1st eigenvector and visualize your results with figures. 
\item Explore the ISOMAP-embedding of the 33 faces on the $k=5$ nearest neighbor graph and compare it against the MDS results. Note: you may try Tenenbaum's Matlab code \\
{\url{https://github.com/yao-lab/yao-lab.github.io/blob/master/data/isomapII.m}}
\item Explore the LLE/MLLE-embedding of the 33 faces on the $k=5$ nearest neighbor graph and compare it against ISOMAP. Note: you may try the following Matlab/Python code \\
{\url{https://github.com/yao-lab/yao-lab.github.io/blob/master/data/lle.m}}, %\\
%{\url{https://scikit-learn.org/stable/modules/manifold.html#modified-locally-linear-embedding}}
\item Explore the LSTA-embedding of the 33 faces on the $k=5$ nearest neighbor graph and compare it against ISOMAP. Note: you may try the following Python code \\
{\url{https://scikit-learn.org/stable/modules/manifold.html#local-tangent-space-alignment}}
\item Explore the 2-D t-SNE embedding of the 33 faces. %Note: you may try the following Python code \\
%{\url{https://scikit-learn.org/stable/modules/manifold.html#t-distributed-stochastic-neighbor-embedding-t-sne}}
\end{enumerate}

You might explore larger datasets with other manifold learning methods below.

\subsection{PubFig dataset} 
The PubFig dataset is at

\subsection{MNIST dataset }

Yann LeCun's website contains original MNIST dataset of 60,000 training images and 10,000 test images.

\url{http://yann.lecun.com/exdb/mnist/}

There are various ways to download and parse MNIST files. For example, Python users may refer to the following website:

\url{https://github.com/datapythonista/mnist}

or MXNET tutorial on mnist

\url{https://mxnet.incubator.apache.org/tutorials/python/mnist.html}

\subsection{Fashion-MNIST dataset}

Zalando's Fashion-MNIST dataset of 60,000 training images and 10,000 test images, of size 28-by-28 in grayscale.

\url{https://github.com/zalandoresearch/fashion-mnist}

%As a reference, here is Jason Wu, Peng Xu, and Nayeon Lee's exploration on the dataset in project 1:

%\url{https://deeplearning-math.github.io/slides/Project1_WuXuLee.pdf}

%\section{Identification of Raphael's paintings from the forgeries}
%
%The following data, provided by Prof. Yang WANG from HKUST,
%
%\url{https://drive.google.com/folderview?id=0B-yDtwSjhaSCZ2FqN3AxQ3NJNTA&usp=sharing}
%
%\noindent contains a 28 digital paintings of Raphael or forgeries. Note that there are both jpeg and tiff files, so be careful with the bit depth in digitization. The following file
%
%\url{https://docs.google.com/document/d/1tMaaSIrYwNFZZ2cEJdx1DfFscIfERd5Dp2U7K1ekjTI/edit}
%
%\noindent contains the labels of such paintings, which are 
%\begin{enumerate}
%\item[1] Maybe Raphael - Disputed
%\item[2] Raphael
%\item[3] Raphael
%\item[4] Raphael
%\item[5] Raphael
%\item[6] Raphael
%\item[7] Maybe Raphael - Disputed
%\item[8] Raphael
%\item[9] Raphael
%\item[10] Maybe Raphael - Disputed
%\item[11] Not Raphael
%\item[12] Not Raphael
%\item[13] Not Raphael
%\item[14] Not Raphael
%\item[15] Not Raphael
%\item[16] Not Raphael
%\item[17] Not Raphael
%\item[18] Not Raphael
%\item[19] Not Raphael
%\item[20] My Drawing (Raphael?)
%\item[21] Raphael
%\item[22] Raphael
%\item[23] Maybe Raphael - Disputed
%\item[24] Raphael
%\item[25] Maybe Raphael - Disputed
%\item[26] Maybe Raphael - Disputed
%\item[27] Raphael
%\item[28] Raphael
%\end{enumerate}
%Can you exploit the known Raphael vs. Not Raphael data to predict the identity of those 6 disputed paintings (maybe Raphael)? The following student poster report seems a good exploration
%
%\url{https://yao-lab.github.io/2015.fall.pku/poster/Raphael_LI%2CYue_1300010601.pdf}
%
%The following paper by Haixia Liu, Raymond Chan, and me studies Van Gogh's paintings which might be a reference for you:
%
%\url{http://dx.doi.org/10.1016/j.acha.2015.11.005}

@@

{\it All the following datasets have been given in project 1. But you may explore these data with various new techniques learned in class, e.g. Spectral clustering and ranking, Transition Path Analysis, Robust PCA via rank-sparsity decomposition or via GAN approach to robust covariance estimate (\url{https://github.com/zhuwzh/Robust-GAN-Scatter}), Manifold Learning, Topological Data Analysis (Mapper, Persistent Homology, etc.). }

\section{Hand-written Digits} The website

\url{http://www-stat.stanford.edu/\~tibs/ElemStatLearn/datasets/zip.digits/}

\noindent contains images of 10 handwritten digits (`$0$',...,`9');

@@

\section{SNPs Data}
This dataset contains a data matrix $X\in \R^{n\times p}$ of about $p=650,000$ columns of SNPs (Single Nucleid Polymorphisms) and $n=1064$ rows of peoples around the world (but there are 21 rows mostly with missing values). Each element is of three choices, $0$ (for `AA'), $1$ (for `AC'), $2$ (for `CC'), and some missing values marked by $9$.

\url{https://drive.google.com/file/d/1KMLPEG91mnzdK2pUlq2BkjOx2BsaZy9s/view?usp=sharing}

\noindent which is big (151MB in zip and 2GB original txt). A fast access in the mainland China can be downloaded from:

\url{https://pan.baidu.com/s/1jrv_UfbwWpi_-x5Rg1XS1A}

\noindent with password {\tt 678e}. Moreover, the following file contains the region where each people comes from, as well as two variables {\texttt{ind1}} and{\texttt{ind2}} such that $X({\texttt{ind1}},{\texttt{ind2}})$ removes all missing values.

\url{https://github.com/yao-lab/yao-lab.github.io/blob/master/data/HGDP_region.mat}

Another cleaned dataset is due to Quanhua MU and Yoonhee Nam:

\begin{itemize}
\item Genotyped data of the 1043 ($n$) subjects. 0(AA), 1(AC), 2(CC). Missing values are removed, only autosomal SNPs were selected ($p\approx 400K$). Google drive link: \\
\url{https://drive.google.com/file/d/1a9I8_akfCMHBRrPMdnWkjyL9fKcQbJJq/view?usp=sharing}
\noindent or 
\url{https://pan.baidu.com/s/1vDi0cLWl6GiWgm7icaZy-w} 
\noindent with password {\tt b5mv}.
\item Sample Information of 1043 subjects. Google drive link: \\
\url{https://drive.google.com/file/d/11Q-8B57WDQnrIV92b-h_WLqDGviiYsm2/view?usp=sharing} \\
\end{itemize}

A good reference for this data can be the following paper in Science,

@@

Explore the genetic variation of those persons with their geographic variations, by MDS/PCA. Since $p$ is big, explore random projections for dimensionality reduction.

\section{Robust PCA: Video Clip}

%\subsection{Video}

The following video clip (shoppingmall) has been widely used in literature for rank-sparsity decomposition of matrices. You may download the Matlab .mat file (50MB) from the following:

\url{https://drive.google.com/file/d/1CuVAG3uWnwq6QmI3vARUizOF01Ubfz9k/view?usp=sharing}

\noindent The original .avi file (234MB) can be downloaded at

\url{https://drive.google.com/file/d/10-wwUl10fzzgvVF_YX0E1bEuU2Q9hGNG/view?usp=sharing}

\noindent For those students in mainland China, another fast access of the data can be found at

\url{https://pan.baidu.com/s/1CNSBhueMLpLiD7gxVpQsOA}

\noindent with access password {\tt z9f6}.

%Dr. HanQin CAI introduced his accelerated alternative projection technique for nonconvex robust matrix decomposition, and his source codes can be downloaded at:
%
%\url{https://github.com/caesarcai/AccAltProj_for_RPCA}

%\subsection{Social Network}
%
%The social network dataset of 100 universities can be downloaded at 
%
%\url{https://escience.rpi.edu/data/DA/fb100/}
%
%\noindent whose readme \url{https://escience.rpi.edu/data/DA/fb100/facebook100_readme_021011.txt}.
%
%Dr. Hanbaek Lyu's online robust dictionary learning algorithms can be found at the following links for your exploration: 
%
%\url{https://github.com/HanbaekLyu/ONMF_ONTF_NDL}
%
%\url{https://github.com/HanbaekLyu/RONMF}
%

@@

When $\theta=\infty$ without noise, classical MDS will work; but for a finite $\theta$ with noisy measurements, SDP approach can be useful. You may try the matlab package SNLSDP by Kim-Chuan Toh, Pratik Biswas, and Yinyu Ye, or the facial reduction speed up by Nathan Krislock and Henry Wolkowicz. Just for your reference, the following version SNLSDP is collected and updated by Mengyue ZHA in the class,

\url{https://github.com/Dolores2333/MATH5473/tree/main/HW5/SNLSDP}

For python users, you may try the Python version of CVX (CVXPY): \url{https://www.cvxpy.org/install/index.html}.

\end{document}
